---
title: "EDS221-day7
format: html
editor: visual
---

## 

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Data Wrangling Refresher

1.  starting from penguins, write a single piped sequence in which you

    1.  only include penguins at dream and bisco islands

    2.  remove the year and sex variables

    3.  add a new colum body_mass_kg with penguin mass converted from g to kg

    4.  rename the siland variable to location

    ```{r}
    dream_biscoe <- penguins %>% 
      filter(island %in% c("Dream", "Biscoe")) %>% 
      select(-year, -sex) %>% 
      mutate(body_mass_kg = body_mass_g / 1000) %>% 
      rename(location = island)
            
    ```

2.  Starting from penguins, write a single piped sequence to

    1.  limit to Adelie penguins

    2.  remove any observations where flipper_length_mm s NA (!is.na() or drop_na())

    3.  group the data by sex

    4.  greate a summary table that contains the mean sd and sample size of flipper lengths for male and female adelie penguins

```{r}
adelie <- penguins %>% filter(species == "Adelie") %>% 
 drop_na(flipper_length_mm, sex) %>% 
  group_by(sex) %>% 
  summarise(mean_flipper = mean(flipper_length_mm), sd_flipper = sd(flipper_length_mm), sample_size = n())
```

Counts by species and island and sex

```{r}
counts_penguins <- penguins %>% count(species, island, sex)
#count includes group_by, summarize, and n() to get counts of observations
```

```{r}
animals <- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
           species = c("bobcat", "coyote", "fox", "squirrel", "bobcat"),
          maturity = c("adult", "juvenile", "adult", "juvenile", "adult")
)
```

```{r}
sites <- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
    full_site_name = c("Goleta Beach","UCSB Lagoon",
                       "Ellwood Mesa","Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
)
```

```{r}
full_join_example <- full_join(animals, sites)
```

```{r}
left_join_example <- left_join(animals, sites)
```

```{r}
inner_join_example <- inner_join(animals, sites)
```

1996-02-18T20:34-07:00

2002-12-05T02:41-07:00
